---
title: "Homework 3"
author: "Alexey Abramov"
date:  "10/6/2020"
output: github_document
---

# Setup
```{r, setup}
library(tidyverse)
library(patchwork)
library(hexbin)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(
  ggthemes::theme_clean() + theme(legend.position = "bottom")
  )

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.colour = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Load the dataset.

```{r}
library(p8105.datasets)
data("instacart")
instacart
```

### Problem 1

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.  

How many aisles, and which are most items from?
There are `nrow(aisles_df)` aisles in this dataset.  
Here we are using the count function and arranging from most to least.

```{r}
aisles_df = instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Let's make a plot.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Make a table with the three most popular items in these specific aisles: baking ingredients, dog food care, and packaged vegetable fruits.

Let's first get these aisles, count them, then rank them.  

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Apples versus ice cream.

If I want average of a product at a certain time, I'm thinking group_by and then summarize functions.  

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour)
```

## Problem 2

1.  Pivot longer, day, activity count in every minute of the day, and check minute of the day in numeric
2.  Add a variable for weekday and then weekend with mutate, and use factors so the day of the week is sensible.  
3.  Variable class
4.  Group_by week and then day, and summarize problem
5.  ggplot with geom_line


## Problem 3

1.  Separate
2.  Count for most commonly observed, rank?
3.  Group_by and summarize, then filter
4.  Then ggplot
5.  Two panel with faceting
6. Two distinct plots with patchwork
7.  tmax and tmin is a scatterplot, but probably better with contour or hexplot
8.  filtering with ridgeplots

```{r}

```

